Percussive Synthesizer Generation with Virtual Listeners

### 2020-08-22 templates_aimusic2020/paper.tex
Updates from Overleaf

To automatically learn from the described features, two groups of synthetic ears are implemented: \emph{two phased ears} (TPEs) and \emph{Mixed Ear Models} (MEMs). We train multiple neural network architectures using different subsets of the FFT features to specialize in \decfirst~ or \decsecond~. Our design of TPE ears allows a combination of these networks to make sequential decisions. To train TPEs for \decfirst, we use all drums in RadarDB and FreeDB (categories do not matter). To learn \enquote{not-drums}, TPEs are given 6000 examples of synthesized noise. For \decsecond, we combine the two databases and merge toms into kicks and rims/shakers into \enquote{other}. We designed and trained a number of neural net architectures with 80\% of this dataset. Using the remaining 20\% of sounds, we achieved 98\% accuracy in \decfirst~and 82\% accuracy in~\decsecond. Our best models both utilize convolutional neural net (CNN) layers that learn from spectrogram data. These accuracy numbers are weak as we did not account for category sizes or cross validate.

@@

The TPE system produces samples in the following categories: \enquote{snare}, \enquote{kick}, \enquote{hat}, \enquote{clap} and \enquote{other} (combination of rims, shakers and unusual percussive sounds). The MEM pipeline does not output the \enquote{other} category, yet the option is available to surveyors when categorizing sounds manually.   
Both authors categorized a subset of the results without knowledge of the assigned labels. Additionally, each responder has the option of labeling samples as \enquote{bad} for samples that they deemed not percussive.

@@

The majority of sounds created by our system are deemed percussive by both surveyors. 30\% of the outputs from the TPE system and 50\% of the MEM system outputs are deemed as non-percussive by at least one person. This leaves much room for improvement with regards to~\decfirst. Our performance with regards to ~\decsecond~ is promising as we achieve moderately high agree-ability scores after dropping \enquote{bad sounds} and \enquote{other}. Despite the drawbacks, the MEM pipeline remains competitive while working with a fraction of the features to learn from and simultaneously making both decisions. Extended survey analysis is provided in Appendix~\ref{appendix:surveys}. Appendix~\ref{appendix:datasets} provides access to generated sounds.

@@

\textbf{Conclusion:} We successfully built a generative systems for creation of percussive sounds via automatic programming of virtual synthesizers. We verified its results with human listeners.
Our work enables not only the creation of new libraries of percussion sounds, but new synthesizer programs which can be modified and studied. 
Manual listening tests revealed much room for improvement, particularly with accurate separation of percussive sounds from the infinitely large set of non-percussive sounds. We had some success in our utilization of latent representations of autoencoder networks as low-dimensional representations of short sound files. \\ 
\textbf{Threats to Validity:} The lack of consistency in training and accuracy measurements makes comparisons between TPEs and MEMs difficult. Many arbitrary design decisions have been made, particularly in the design of our TPE classifiers and selection of datasets for training and testing. We cannot quantify the novelty of our results. Our requirement for training data remains high. \\
\textbf{Future Work:} Effective implementation of few-shot learning is a priority. Program generation can be improved by reinforcement learning and other heuristics.

@@

\subsubsection{Model Selection}
Using our encoding and envelope features to represent audio, five classification models were trained for the task of categorizing the five different sound groups. For hyper-parameter optimization, the models were trained using 5-fold cross validation and 80/20 train-to-test ratio. The F-Score result of each cross-validation is the unweighted average F-Score of all groups. For inter-model comparisons, the procedure is the same except 10-fold cross validations are used. Our models were derived from scikit-learn's implementations of these classifiers~\cite{pedregosa2011scikit}. Before inter-model comparisons, we conducted a grid-search for each model on at least one of its possibly decisive hyper-parameters. The classifiers and other notable specifications are presented in Table~\ref{table:mem_model_selection}. 
\begin{table}[t]
    \centering \hspace*{-0.8cm}
    \begin{threeparttable}
    \begin{tabular}[width=0.95\paperwidth]{|l|l|l|}
    \hline
    Model name & Tuned Parameters\tnote{\dag}  & Used Weights? \tnote{\ddag} \\\hline
     Support Vector Classifier (SVC) &  Gamma:0.001, C:100, kernel:rbf & Yes\\
     LinearSVC & C:10 & Yes\\
     K Nearest Neighbors & Num. Neighbors:30 &  No \\
     Random Forest Classifier & Num Estimators:500 & Yes \\
     Extra Trees Classifier & Num Estimators:1100 & Yes\\
     \hline
    \end{tabular}
    \caption{Models implemented for comparison using envelope and embedded features. }
    \begin{tablenotes}
    \item[\dag] Tuned parameters values are based on grid-searching for best f-score. Parameters not mentioned have neither been tuned nor changed from scikit-learn's default values (as of version 0.23)
    \item[\ddag] Class weights are used unless not applicable to classifier.
    \end{tablenotes}
    \label{table:mem_model_selection}
    \end{threeparttable}
\end{table}
\begin{table}[htbp]
\centering
\begin{tabular}{|l|l|l|l|l|l|}
\hline
 DB Name & kick & snare & clap & hat & Synthetic Noise\\\hline
 MixedEarDB & 1334 & 1035 & 401 & 1275 & 1000 \\ \hline
\end{tabular}
\caption{A database put together by combination of radarDB, FreeDB and NoiseDB}
\label{db:memDB}
\end{table}

>>>>

To automatically learn from the described features, two groups of synthetic ears are implemented: \emph{two phased ears} (TPEs) and \emph{Mixed Ear Models} (MEMs). We train multiple neural network architectures using different subsets of the FFT features to specialize in \decfirst~or \decsecond~. Our design of TPE ears allows a combination of these networks to make sequential decisions. To train TPEs for \decfirst, we use all drums in RadarDB and FreeDB (categories do not matter). To learn \enquote{not-drums}, TPEs are given 6000 examples of synthesized noise. For \decsecond, we combine the two databases and merge toms into kicks and rims/shakers into \enquote{other}. We designed and trained a number of neural net architectures with 80\% of this dataset. Using the remaining 20\% of sounds, we achieved 98\% accuracy in \decfirst~and 82\% accuracy in~\decsecond. Our best models both utilize convolutional neural net (CNN) layers that learn from spectrogram data. These accuracy numbers are weak as we did not account for category sizes or cross validate.

@@

The TPE system produces samples in the following categories: \enquote{snare}, \enquote{kick}, \enquote{hat}, \enquote{clap} and \enquote{other} (combination of rims, shakers and unusual percussive sounds). The MEM system does not output the \enquote{other} category, yet the option is available to surveyors when categorizing sounds manually.   
Both authors categorized a subset of the results without knowledge of the assigned labels. Additionally, each responder has the option of labeling samples as \enquote{bad} for samples that they deemed not percussive.

@@

The majority of sounds created by our system are deemed percussive by both surveyors. 30\% of the outputs from the TPE system and 50\% of the MEM system outputs are deemed as non-percussive by at least one person. This leaves much room for improvement with regards to~\decfirst. Our performance with regards to \decsecond~is promising as we achieve moderately high agree-ability scores after dropping \enquote{bad sounds} and \enquote{other}. Despite the drawbacks, the MEM pipeline remains competitive while working with a fraction of the features to learn from and simultaneously making both decisions. Extended survey analysis is provided in Appendix~\ref{appendix:surveys}. Appendix~\ref{appendix:datasets} provides access to generated sounds.

@@

\textbf{Conclusion:} We successfully built a generative system for creation of percussive sounds via automatic programming of virtual synthesizers. We verified its results with human listeners.
Our work enables not only the creation of new libraries of percussion sounds, but new synthesizer programs which can be modified and studied. 
Manual listening tests revealed much room for improvement, particularly with accurate separation of percussive sounds from the infinitely large set of non-percussive sounds. We had some success in our utilization of latent representations of autoencoder networks as low-dimensional representations of short sound files. \\ 
\textbf{Threats to Validity:} The lack of consistency in training and accuracy measurements makes comparisons between TPEs and MEMs difficult. Many arbitrary design decisions have been made, particularly in the design of our TPE classifiers and selection of datasets for training and testing. We cannot quantify the novelty of our results. Our requirement for training data remains high. \\
\textbf{Future Work:} Effective implementation of few-shot learning is a priority. Program generation can be improved by reinforcement learning and other heuristics.

@@

\subsubsection{Model Selection}
Using our encoding and envelope features to represent audio, five classification models were trained for the task of categorizing the five different sound groups. For hyper-parameter optimization, the models were trained using 5-fold cross validation and 80/20 train-to-test ratio. The F-Score result of each cross-validation is the unweighted average F-Score of all groups. For inter-model comparisons, the procedure is the same except 10-fold cross validations are used. Our models were derived from scikit-learn's implementations of these classifiers~\cite{pedregosa2011scikit}. Before inter-model comparisons, we conducted a grid-search for each model on at least one of its possibly decisive hyper-parameters. The classifiers and other notable specifications are presented in Table~\ref{table:mem_model_selection}. 
\begin{table}[t]
    \centering \hspace*{-0.8cm}
    \begin{threeparttable}
    \begin{tabular}[width=0.95\paperwidth]{|l|l|l|}
    \hline
    Model name & Tuned Parameters\tnote{\dag}  & Used Weights? \tnote{\ddag} \\\hline
     Support Vector Classifier (SVC) &  Gamma:0.001, C:100, kernel:rbf & Yes\\
     LinearSVC & C:10 & Yes\\
     K Nearest Neighbors & Num. Neighbors:30 &  No \\
     Random Forest Classifier & Num Estimators:500 & Yes \\
     Extra Trees Classifier & Num Estimators:1100 & Yes\\
     \hline
    \end{tabular}
    \caption{Models implemented for comparison using envelope and embedded features. }
    \begin{tablenotes}
    \item[\dag] Tuned parameters values are based on grid-searching for best f-score. Parameters not mentioned have neither been tuned nor changed from scikit-learn's default values (as of version 0.23)
    \item[\ddag] Class weights are used unless not applicable to classifier.
    \end{tablenotes}
    \label{table:mem_model_selection}
    \end{threeparttable}
\end{table}
\begin{table}[htbp]
\centering
\begin{tabular}{|l|l|l|l|l|l|}
\hline
 DB Name & kick & snare & clap & hat & Synthetic Noise\\\hline
 MixedEarDB & 1334 & 1035 & 401 & 1275 & 1000 \\ \hline
\end{tabular}
\caption{A database put together by combination of RadarDB, FreeDB and NoiseDB}
\label{db:memDB}
\end{table}
